---
title: "Decision Trees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decision Trees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(openqaly)
```

## Introduction

Decision trees model one-time clinical events where patients follow different
pathways with assigned probabilities and payoffs. In health economics, they are
used for short-term clinical decisions such as treatment response, surgical
outcomes, or diagnostic testing.

openqaly supports two modes of decision tree modeling:

1. **Standalone decision tree models** — single-point-in-time models where all
   outcomes and costs are captured in the tree itself.
2. **Decision trees preceding a Markov or PSM model** — the tree determines
   initial state allocations and one-time costs, then a Markov or PSM model
   simulates long-term outcomes.

This vignette demonstrates both modes, tree visualization, and the probability
query system used in value formulas.

---

## Example 1: Standalone Decision Tree Model

### Motivation

A simple drug comparison: two drugs with different response rates. Responders
incur higher treatment costs but gain more QALYs. Non-responders receive
palliative care at lower cost and lower QALY gain.

| Outcome | Cost | QALYs |
|---------|------|-------|
| Respond | 15,000 | 0.9 |
| No respond | 5,000 | 0.5 |

Drug A has a 70% response rate; Drug B has 40%.

### Building the Model

```{r dt-standalone-model}
dt_model <- define_model("decision_tree") |>
  # Payoff variables
  add_variable("cost_respond", 15000) |>
  add_variable("cost_no_respond", 5000) |>
  add_variable("qaly_respond", 0.9) |>
  add_variable("qaly_no_respond", 0.5) |>
  # Strategy-specific response rates
  add_variable("p_respond", 0.7, strategy = "drug_a") |>
  add_variable("p_respond", 0.4, strategy = "drug_b") |>
  # Strategies
  add_strategy("drug_a", display_name = "Drug A") |>
  add_strategy("drug_b", display_name = "Drug B") |>
  # Decision tree structure
  add_tree_node("drug_choice", "root", parent = NA, formula = 1) |>
  add_tree_node("drug_choice", "respond", parent = "root",
                formula = p_respond) |>
  add_tree_node("drug_choice", "no_respond", parent = "root",
                formula = C) |>
  # Configure the decision tree (duration = 0 for standalone)
  set_decision_tree("drug_choice", duration = 0, duration_unit = "days") |>
  # Values — use p() to weight payoffs by branch probability
  add_value("cost", type = "cost", state = "decision_tree",
            formula = p(respond, drug_choice) * cost_respond +
                      p(no_respond, drug_choice) * cost_no_respond) |>
  add_value("qaly", type = "outcome", state = "decision_tree",
            formula = p(respond, drug_choice) * qaly_respond +
                      p(no_respond, drug_choice) * qaly_no_respond) |>
  # Summaries
  add_summary("total_cost", values = "cost",
              display_name = "Total Cost", type = "cost") |>
  add_summary("total_qaly", values = "qaly",
              display_name = "QALYs", type = "outcome")
```

Key points:

- `define_model("decision_tree")` creates a standalone DT model (no states or
  transitions)
- `add_tree_node()` builds the tree structure — each node has a name, parent,
  and probability formula
- `C` computes the complement probability (1 minus the sum of sibling
  probabilities)
- `set_decision_tree()` activates the tree with `duration = 0` for standalone
  models
- Values use `state = "decision_tree"` and reference branch probabilities via
  `p()`

### Running the Model

```{r dt-standalone-run}
dt_results <- run_model(dt_model)
```

### Viewing Results

```{r dt-standalone-table}
outcomes_table(dt_results, outcome = "total_qaly")
```

```{r dt-standalone-plot, fig.width=9, fig.height=6}
outcomes_plot_bar(dt_results, outcome = "total_qaly")
```

---

## Example 2: Decision Tree in Front of a Markov Model

### Motivation

A surgery with two possible outcomes determines long-term health trajectories.
The surgical outcome feeds into a Markov model where patients transition between
health states over time.

| Phase | Component | Description |
|-------|-----------|-------------|
| DT phase | Surgery | Success (→ healthy) or failure (→ dead) |
| Markov phase | Long-term | Healthy → Dead transition over time |

Two strategies are compared: aggressive surgery (90% success, higher cost) vs
conservative surgery (70% success, lower cost).

### Building the Model

```{r dt-markov-model}
markov_dt_model <- define_model("markov") |>
  # Settings for the Markov phase
  set_settings(
    timeframe = 10,
    timeframe_unit = "years",
    cycle_length = 1,
    cycle_length_unit = "years",
    discount_cost = 3.5,
    discount_outcomes = 3.5
  ) |>
  # Strategies
  add_strategy("aggressive", display_name = "Aggressive Surgery") |>
  add_strategy("conservative", display_name = "Conservative Surgery") |>
  # Strategy-specific success rates
  add_variable("p_success", 0.9, strategy = "aggressive") |>
  add_variable("p_success", 0.7, strategy = "conservative") |>
  # Cost variables
  add_variable("c_surgery_success", 50000) |>
  add_variable("c_surgery_failure", 75000) |>
  add_variable("c_treatment", 1000) |>
  # Decision tree for surgery outcomes
  add_tree_node("surgery", "root", parent = NA, formula = 1) |>
  add_tree_node("surgery", "success", parent = "root",
                formula = p_success) |>
  add_tree_node("surgery", "failure", parent = "root", formula = C) |>
  # DT duration offsets Markov discounting by 30 days
  set_decision_tree("surgery", duration = 30, duration_unit = "days") |>
  # Markov states — initial probabilities reference DT outcomes
  add_state("healthy", initial_prob = "p(success, surgery)") |>
  add_state("dead", initial_prob = "C") |>
  # Transitions
  add_transition("healthy", "dead", 0.1) |>
  add_transition("healthy", "healthy", C) |>
  add_transition("dead", "dead", 1) |>
  # DT-phase value: one-time surgery cost (not discounted)
  add_value("surgery_cost", type = "cost", state = "decision_tree",
            formula = p(success, surgery) * c_surgery_success +
                      p(failure, surgery) * c_surgery_failure) |>
  # Markov-phase values: ongoing costs and QALYs (discounted with DT offset)
  add_value("treatment_cost", type = "cost", state = "healthy",
            formula = c_treatment) |>
  add_value("qaly", type = "outcome", state = "healthy", formula = 1) |>
  # Summaries
  add_summary("total_cost", type = "cost",
              values = "surgery_cost,treatment_cost",
              display_name = "Total Cost") |>
  add_summary("total_qaly", type = "outcome",
              values = "qaly",
              display_name = "QALYs")
```

Key differences from standalone DT:

- `define_model("markov")` — the base model type is Markov, not decision_tree
- `set_decision_tree(duration = 30, duration_unit = "days")` — non-zero duration
  creates a discounting offset for the Markov phase
- `initial_prob = "p(success, surgery)"` — tree outcomes set initial state
  probabilities
- DT values (`state = "decision_tree"`) are one-time, undiscounted payoffs
- Markov values (`state = "healthy"`) accrue per-cycle and are discounted
  starting after the 30-day DT offset

### Running the Model

```{r dt-markov-run}
markov_dt_results <- run_model(markov_dt_model)
```

### Viewing Results

```{r dt-markov-trace, fig.width=9, fig.height=6}
# Markov phase state occupancy
trace_plot_area(markov_dt_results)
```

```{r dt-markov-outcomes}
outcomes_table(markov_dt_results, outcome = "total_qaly")
```

---

## Plotting Decision Trees

### Tree Structure Plot

`plot_decision_tree()` visualizes the tree structure from a model object (before
running). It shows node shapes (circles for chance nodes, triangles for terminal
nodes) and formula labels on edges.

```{r dt-structure-plot, fig.width=9, fig.height=6}
plot_decision_tree(dt_model)
```

For models with multiple trees, specify which tree to plot:

```r
plot_decision_tree(model, tree_name = "my_tree")
```

### Evaluated Tree Sankey Plot

`plot_decision_tree_sankey()` visualizes the evaluated tree from `run_model()`
output. Flow widths are proportional to computed probabilities, and flows are
colored by the first branch.

```{r dt-sankey-plot, fig.width=9, fig.height=6}
plot_decision_tree_sankey(dt_results)
```

The Sankey automatically facets when there are multiple strategies or groups.
You can filter with the `strategies` and `groups` parameters:

```r
plot_decision_tree_sankey(results, strategies = "drug_a")
```

Note: Sankey plots require base-case results — they are not available for PSA
or DSA output.

---

## Querying Probabilities with `p()`

The `p()` function is the primary interface for extracting probabilities from
evaluated decision trees. It supports set operations, conditional probabilities,
and is used inside `add_value()` formulas to weight payoffs by branch
probability.

### Example Tree for This Section

We build a more detailed tree to demonstrate all `p()` operators:

```
trial_tree
├── treatment (0.5)
│   ├── response (0.7) [good_outcome]
│   │   ├── cured (0.4)
│   │   └── managed (C = 0.6)
│   └── no_response (C = 0.3) [bad_outcome]
└── control (C = 0.5)
    ├── response_ctrl (0.5) [good_outcome]
    └── no_response_ctrl (C = 0.5) [bad_outcome]
```

Terminal node joint probabilities:

| Terminal | Probability | Tags |
|----------|-------------|------|
| cured | 0.5 × 0.7 × 0.4 = 0.14 | treatment, response, good_outcome |
| managed | 0.5 × 0.7 × 0.6 = 0.21 | treatment, response, good_outcome |
| no_response | 0.5 × 0.3 = 0.15 | treatment, bad_outcome |
| response_ctrl | 0.5 × 0.5 = 0.25 | control, good_outcome |
| no_response_ctrl | 0.5 × 0.5 = 0.25 | control, bad_outcome |

```{r p-demo-tree}
p_demo <- define_model("decision_tree") |>
  add_strategy("s") |>
  add_tree_node("trial_tree", "root", parent = NA, formula = 1) |>
  add_tree_node("trial_tree", "treatment", parent = "root",
                formula = 0.5) |>
  add_tree_node("trial_tree", "control", parent = "root", formula = C) |>
  add_tree_node("trial_tree", "response", parent = "treatment",
                formula = 0.7, tags = "good_outcome") |>
  add_tree_node("trial_tree", "no_response", parent = "treatment",
                formula = C, tags = "bad_outcome") |>
  add_tree_node("trial_tree", "cured", parent = "response",
                formula = 0.4) |>
  add_tree_node("trial_tree", "managed", parent = "response",
                formula = C) |>
  add_tree_node("trial_tree", "response_ctrl", parent = "control",
                formula = 0.5, tags = "good_outcome") |>
  add_tree_node("trial_tree", "no_response_ctrl", parent = "control",
                formula = C, tags = "bad_outcome")

# Evaluate the tree directly for interactive demonstration
trial_tree <- decision_tree(p_demo$trees, "trial_tree", 0)
```

### How `p()` Works

`p()` takes two arguments: an **unevaluated expression** specifying which nodes
to query, and the **evaluated tree object**. Inside `p()`:

1. Each node name and tag name becomes a **set of terminal nodes**
2. A node name resolves to all terminal nodes reachable through that node
3. A tag name resolves to all terminal nodes carrying that tag
4. Operators combine these sets
5. The joint probabilities of the resulting terminal nodes are summed

### Basic Queries: Node Probabilities

`p(node_name, tree)` returns the sum of joint probabilities for all terminal
nodes reachable through that node:

```{r p-basic}
# Single terminal node — joint probability (0.5 * 0.7 * 0.4 = 0.14)
p(cured, trial_tree)

# Parent node — sum of descendant terminal probabilities
# response includes: cured (0.14) + managed (0.21) = 0.35
p(response, trial_tree)

# Top-level branch (0.50)
p(treatment, trial_tree)

# Root — always 1.0 (all terminal nodes)
p(root, trial_tree)
```

### Tags and Named Groups

The `tags` parameter on `add_tree_node()` assigns custom tags to nodes. A tag
on a non-terminal node propagates to all its terminal descendants. Each node's
name is also automatically a tag for itself and its descendants.

Custom tags group nodes across branches:

```{r p-tags}
# "good_outcome" tag spans both arms:
# cured (0.14) + managed (0.21) + response_ctrl (0.25) = 0.60
p(good_outcome, trial_tree)

# "bad_outcome" tag:
# no_response (0.15) + no_response_ctrl (0.25) = 0.40
p(bad_outcome, trial_tree)
```

### The `%and%` Operator (Intersection)

`p(A %and% B, tree)` returns the probability of terminal nodes that belong
to **both** set A and set B:

```{r p-and}
# Terminal nodes under "treatment" AND tagged "good_outcome"
# = cured (0.14) + managed (0.21) = 0.35
p(treatment %and% good_outcome, trial_tree)
```

### The `%or%` Operator (Union)

`p(A %or% B, tree)` returns the probability of terminal nodes in A **or** B
(or both):

```{r p-or}
# Terminals in treatment OR good_outcome
# = treatment(0.50) + good_outcome(0.60) - intersection(0.35) = 0.75
p(treatment %or% good_outcome, trial_tree)
```

### The `-` Operator (Complement)

`p(-A, tree)` returns the probability of all terminal nodes **not** in set A:

```{r p-not}
# All terminals NOT under treatment = control branch (0.50)
p(-treatment, trial_tree)

# All terminals NOT tagged good_outcome (0.40)
p(-good_outcome, trial_tree)
```

### The `|` Operator (Conditional Probability)

`p(A | B, tree)` computes the conditional probability P(A|B) = P(A ∩ B) / P(B):

```{r p-given}
# Among treatment terminals, what fraction are good outcomes?
# = P(treatment AND good_outcome) / P(treatment) = 0.35 / 0.50 = 0.70
p(good_outcome | treatment, trial_tree)

# Among good outcome terminals, what fraction are from treatment?
# = P(treatment AND good_outcome) / P(good_outcome) = 0.35 / 0.60
p(treatment | good_outcome, trial_tree)
```

### Combining Operators

Operators can be combined for complex queries:

```{r p-combined}
# Bad outcomes within treatment branch
# = no_response = 0.15
p(-good_outcome %and% treatment, trial_tree)

# Treatment given not-bad-outcome
# = P(treatment AND good_outcome) / P(good_outcome) = 0.35 / 0.60
p(treatment | -bad_outcome, trial_tree)
```

### Using `p()` in Value Formulas

In model building, `p()` is used inside `add_value()` formulas to weight
payoffs by branch probability. The tree variable name in the formula references
the evaluated tree in the model's namespace:

```r
add_value("cost", type = "cost", state = "decision_tree",
          formula = p(respond, my_tree) * treatment_cost +
                    p(no_respond, my_tree) * palliative_cost)
```

---

## Key Concepts

### The Complement Operator `C`

`C` calculates `1 - sum(sibling_probabilities)` for a tree node. It ensures
branch probabilities sum to 1 at each level. Only one `C` is allowed per parent
node.

```r
add_tree_node("tree", "node_a", parent = "root", formula = 0.6) |>
add_tree_node("tree", "node_b", parent = "root", formula = C)  # = 0.4
```

Note: `C` in tree node formulas (complement of siblings) is different from the
`-` operator in `p()` queries (complement of terminal node sets).

### Duration and Discounting

`set_decision_tree(duration, duration_unit)` specifies how long the DT phase
lasts. This affects the downstream model:

- **Duration = 0** (standalone DT): No discounting applied to DT values.
- **Duration > 0** (DT + Markov/PSM): The Markov/PSM cycle count is reduced by
  the DT duration, and discounting of state values is offset by the DT duration.
- DT-phase values (`state = "decision_tree"`) are **always undiscounted** by
  default, regardless of duration.

---

## Comparison

| Feature | Standalone DT | DT + Markov | DT + PSM |
|---------|--------------|-------------|----------|
| Model type | `"decision_tree"` | `"markov"` | `"psm"` |
| States | None | Required | Required |
| Time dimension | Single point | Cycles | Cycles |
| DT duration | 0 (typical) | Non-zero | Non-zero |
| DT values | One-time payoffs | One-time payoffs | One-time payoffs |
| State values | Not applicable | Per-cycle | Per-cycle |
| Discounting | None by default | Offset by DT duration | Offset by DT duration |
