---
title: "Setting Up Probabilistic Sensitivity Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting Up Probabilistic Sensitivity Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette covers how to configure sampling distributions for probabilistic sensitivity analysis (PSA) in openqaly. For visualizing and interpreting PSA results, see the companion vignette `vignette("psa_results")`.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
# Skip on CI and CRAN (R CMD check sets _R_CHECK_PACKAGE_NAME_)
if (identical(Sys.getenv("CI"), "true") || nzchar(Sys.getenv("_R_CHECK_PACKAGE_NAME_"))) {
  message("Skipping vignette build on CI/CRAN")
  knitr::knit_exit()
}

library(openqaly)
```

## Building a Model with PSA

We'll build a Markov model that demonstrates every available sampling distribution type. The model represents a chronic disease with three health states (stable, progressed, dead) and two treatment strategies.

```{r model}
model <- define_model("markov") |>
  set_settings(
    n_cycles = 20,
    cycle_length = 1,
    cycle_length_unit = "years",
    discount_cost = 3.5,
    discount_outcomes = 3.5
  ) |>
  add_strategy("standard") |>
  add_strategy("novel") |>
  add_state("stable", initial_prob = 1) |>
  add_state("progressed", initial_prob = 0) |>
  add_state("dead", initial_prob = 0)
```

## Univariate Distributions

Univariate sampling distributions are specified via the `sampling` argument in `add_variable()`. The special keyword `bc` references the variable's base case value, allowing relative specifications.

### Beta distribution

For probabilities bounded between 0 and 1. Can be specified with (mean, sd) or (shape1, shape2).

```{r beta}
model <- model |>
  add_variable("p_progression_death", 0.15, sampling = beta(bc, bc * 0.25)) |>
  add_variable("p_adverse_event", 0.08, sampling = beta(shape1 = 8, shape2 = 92))
```

### Normal distribution

For continuous parameters without natural bounds.

```{r normal}
model <- model |>
  add_variable("treatment_effect", 0.5, strategy = "novel",
               sampling = normal(bc, 0.1))
```

### Lognormal distribution

For parameters that must be positive (e.g., costs, hazard ratios). Can be parameterized on the natural or log scale.

```{r lognormal}
model <- model |>
  add_variable("cost_progressed", 12000,
               sampling = lognormal(mean = bc, sd = bc * 0.25))
```

### Gamma distribution

For positive continuous parameters. Common for costs and resource use.

```{r gamma}
model <- model |>
  add_variable("cost_stable", 3000, sampling = gamma(mean = bc, sd = bc * 0.2)) |>
  add_variable("cost_treatment", 8000, strategy = "novel",
               sampling = gamma(mean = bc, sd = bc * 0.15))
```

### Uniform distribution

For parameters with known bounds but no other distributional information.

```{r uniform}
model <- model |>
  add_variable("cost_treatment", 1000, strategy = "standard",
               sampling = uniform(min = 500, max = 1500))
```

### Triangular distribution

When only minimum, most likely (mode), and maximum values are available, typically from expert opinion.

```{r triangular}
model <- model |>
  add_variable("u_progressed", 0.55,
               sampling = triangular(min = 0.35, mode = 0.55, max = 0.75))
```

## Multivariate Distributions

Use `add_multivariate_sampling()` when parameters need to be sampled jointly. This preserves correlations between parameters across PSA simulations.

### Dirichlet distribution

For transition probabilities that must sum to 1. Specify `n` (effective sample size) and alpha values are derived automatically as `n * base_case_value`. Higher `n` means less uncertainty.

```{r dirichlet}
model <- model |>
  add_variable("p_stable_stable", 0.75) |>
  add_variable("p_stable_progressed", 0.20) |>
  add_variable("p_stable_dead_extra", 0.05) |>
  add_multivariate_sampling(
    name = "stable_transitions",
    type = "dirichlet",
    variables = c("p_stable_stable", "p_stable_progressed", "p_stable_dead_extra"),
    n = 200
  )
```

### Multivariate normal

For correlated continuous parameters. The covariance matrix is stored as a model table via `add_table()` and referenced by name. Mean values are taken automatically from the variables' base case values.

This is the standard approach when working with fitted model outputs (e.g., from flexsurv), where `vcov()` provides the variance-covariance matrix directly.

```{r mvnormal}
# Three correlated utility parameters
# sd: stable=0.05, progressed=0.08, dead=0 (fixed)
# cor(stable, progressed) = 0.6
# Only stable and progressed are uncertain
cov_matrix <- data.frame(
  u_stable    = c(0.05^2,          0.6 * 0.05 * 0.08, 0.3 * 0.05 * 0.02),
  u_treatment = c(0.6 * 0.05 * 0.08, 0.08^2,          0.4 * 0.08 * 0.02),
  u_benefit   = c(0.3 * 0.05 * 0.02, 0.4 * 0.08 * 0.02, 0.02^2)
)

model <- model |>
  add_variable("u_stable", 0.85) |>
  add_variable("u_treatment", 0.10) |>
  add_variable("u_benefit", 0.05) |>
  add_table("utility_cov", cov_matrix) |>
  add_multivariate_sampling(
    name = "correlated_utilities",
    type = "mvnormal",
    variables = c("u_stable", "u_treatment", "u_benefit"),
    covariance = "utility_cov"
  )
```

In practice, the covariance matrix often comes directly from a fitted model:

```{r flexsurv_example, eval = FALSE}
# Example with flexsurv (not executed)
# fit <- flexsurvreg(Surv(time, status) ~ 1, data = dat, dist = "gengamma")
model |>
  add_variable("mu", coef(fit)["mu"]) |>
  add_variable("sigma", coef(fit)["sigma"]) |>
  add_variable("Q", coef(fit)["Q"]) |>
  add_table("surv_vcov", as.data.frame(vcov(fit))) |>
  add_multivariate_sampling(
    name = "survival_params",
    type = "mvnormal",
    variables = c("mu", "sigma", "Q"),
    covariance = "surv_vcov"
  )
```

### Multinomial distribution

For categorical outcomes where exactly one category is selected per simulation (one-hot encoded). Probabilities are taken from the variables' base case values.

```{r multinomial}
model <- model |>
  add_variable("start_early", 0.6) |>
  add_variable("start_mid", 0.3) |>
  add_variable("start_late", 0.1) |>
  add_multivariate_sampling(
    name = "disease_stage_at_entry",
    type = "multinomial",
    variables = c("start_early", "start_mid", "start_late"),
    size = 1
  )
```

## Strategy-Specific Sampling

Multivariate sampling can be restricted to a specific strategy using the `strategy` parameter. The sampling only applies when running that strategy's segment; other strategies use the base case values.

```{r strategy_specific}
# Novel strategy has correlated efficacy parameters
novel_cov <- data.frame(
  efficacy_response = c(0.01,   0.003,  0.002),
  efficacy_duration = c(0.003,  0.04,   0.005),
  efficacy_qol      = c(0.002,  0.005,  0.0025)
)

model <- model |>
  add_variable("efficacy_response", 0.6) |>
  add_variable("efficacy_duration", 1.5) |>
  add_variable("efficacy_qol", 0.08) |>
  add_table("novel_efficacy_cov", novel_cov) |>
  add_multivariate_sampling(
    name = "novel_efficacy",
    type = "mvnormal",
    variables = c("efficacy_response", "efficacy_duration", "efficacy_qol"),
    strategy = "novel",
    covariance = "novel_efficacy_cov"
  )
```

## Completing and Running the Model

Now we add transitions, values, and summaries to complete the model, then run the PSA.

```{r complete_model}
model <- model |>
  # Transitions
  add_transition("stable", "stable", "p_stable_stable") |>
  add_transition("stable", "progressed", "p_stable_progressed") |>
  add_transition("stable", "dead", "p_stable_dead_extra") |>
  add_transition("progressed", "progressed", "1 - p_progression_death") |>
  add_transition("progressed", "dead", "p_progression_death") |>
  add_transition("dead", "dead", "1") |>
  # Values
  add_value("cost", "cost_stable + cost_treatment", state = "stable", type = "cost") |>
  add_value("cost", "cost_progressed", state = "progressed", type = "cost") |>
  add_value("cost", "0", state = "dead", type = "cost") |>
  add_value("qaly", "u_stable", state = "stable", type = "outcome") |>
  add_value("qaly", "u_progressed", state = "progressed", type = "outcome") |>
  add_value("qaly", "0", state = "dead", type = "outcome") |>
  # Summaries
  add_summary("total_cost", "cost", type = "cost") |>
  add_summary("total_qaly", "qaly", type = "outcome")
```

```{r run_psa}
psa_results <- run_psa(model, n_sim = 500, seed = 42)
```

The PSA samples all configured distributions simultaneously across `n_sim` iterations. Univariate distributions are sampled independently; multivariate distributions preserve the specified correlation structure.

## Verifying Multivariate Correlation Structure

Use `psa_parameter_scatter_matrix()` to verify that the sampled parameters exhibit the expected correlation structure. This is especially important for multivariate distributions.

### Dirichlet: transition probabilities

The Dirichlet distribution ensures the three transition probabilities from the stable state sum to 1. The scatter matrix shows the negative correlations induced by this constraint.

```{r scatter_dirichlet, fig.width=9, fig.height=9}
psa_parameter_scatter_matrix(
  psa_results,
  variables = c("p_stable_stable", "p_stable_progressed", "p_stable_dead_extra"),
  strategies = "standard"
)
```

### Multivariate normal: correlated utilities

The scatter matrix confirms the correlation structure specified by the covariance table.

```{r scatter_mvnormal, fig.width=9, fig.height=9}
psa_parameter_scatter_matrix(
  psa_results,
  variables = c("u_stable", "u_treatment", "u_benefit"),
  strategies = "standard"
)
```

### Strategy-specific multivariate normal: novel efficacy

Since `novel_efficacy` targets the "novel" strategy, these correlations only appear for that strategy.

```{r scatter_strategy, fig.width=9, fig.height=9}
psa_parameter_scatter_matrix(
  psa_results,
  variables = c("efficacy_response", "efficacy_duration", "efficacy_qol"),
  strategies = "novel"
)
```

### Multinomial: disease stage at entry

Multinomial sampling produces discrete one-hot outcomes (0 or 1). Use `psa_parameter_pair_plot()` instead of the scatter matrix — it automatically detects discrete variables and uses bar charts and dot plots.

```{r pair_multinomial, fig.width=9, fig.height=9}
psa_parameter_pair_plot(
  psa_results,
  variables = c("start_early", "start_mid", "start_late"),
  strategies = "standard"
)
```

See `vignette("psa_results")` for the full range of PSA visualization options.
