---
title: "Probabilistic Sensitivity Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Probabilistic Sensitivity Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to configure, run, and visualize probabilistic sensitivity analysis (PSA) using openqaly. PSA simultaneously varies all uncertain parameters according to their probability distributions to characterize overall uncertainty in model results. The examples use a sample model included with the package.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(openqaly)

model <- read_model(system.file("models", "example_markov", package = "openqaly"))
```

```{r psa-config, include = FALSE}
# Configure PSA sampling distributions for the examples
# Update the sampling column directly for existing variables
model$variables <- model$variables %>%
  dplyr::mutate(sampling = dplyr::case_when(
    name == "p_mild" ~ "beta(bc, bc * 0.2)",
    name == "p_severe" ~ "beta(bc, bc * 0.2)",
    name == "p_death_severe" ~ "beta(bc, bc * 0.25)",
    name == "c_mild" ~ "gamma(bc, bc * 0.2)",
    name == "c_severe" ~ "gamma(bc, bc * 0.2)",
    name == "u_mild" ~ "beta(bc, bc * 0.1)",
    name == "u_severe" ~ "beta(bc, bc * 0.15)",
    TRUE ~ sampling
  ))
```

## Configuring PSA Sampling Distributions

Before running PSA, you must specify probability distributions for uncertain parameters using the `sampling` argument in `add_variable()`.

### Basic usage with a normal distribution

```{r, eval = FALSE}
model <- model %>%
  add_variable("c_severe", 8000, sampling = normal(bc, 1000))
```

### Using the `bc` keyword for relative specifications

The `bc` keyword references the base case value, allowing relative distribution specifications.

```{r, eval = FALSE}
model <- model %>%
  add_variable("p_mild", 0.15, sampling = normal(bc, bc * 0.1))
```

### Different distribution types

Common distributions for health economic parameters:

```{r, eval = FALSE}
# Beta distribution for probabilities (constrained to 0-1)
model <- model %>%
  add_variable("p_death", 0.02, sampling = beta(bc, bc * 0.2))

# Lognormal distribution for costs (positive values only)
model <- model %>%
  add_variable("c_treatment", 5000, sampling = lognormal(bc, bc * 0.3))

# Gamma distribution for resource use
model <- model %>%
  add_variable("n_visits", 12, sampling = gamma(bc, bc * 0.25))
```

### Multivariate sampling for correlated parameters

Use `add_multivariate_sampling()` when parameters should be sampled together with correlation.

```{r, eval = FALSE}
model <- model %>%
  add_multivariate_sampling(
    name = "utility_correlation",
    distribution = multi_normal(
      means = c(0.8, 0.6),
      covariance = matrix(c(0.01, 0.005, 0.005, 0.02), nrow = 2)
    ),
    variables = c("u_mild", "u_severe")
  )
```

## Running PSA

```{r, include = FALSE}
# Run PSA for the examples
psa_results <- run_psa(model, n_sim = 1000, seed = 123)
```

The `run_psa()` function executes the model for each Monte Carlo simulation, sampling parameters from their specified distributions.

```{r, eval = FALSE}
psa_results <- run_psa(model, n_sim = 1000, seed = 123)
```

## PSA Scatter Plots

Scatter plots visualize simulation results on the cost-effectiveness plane.

### Cost-Effectiveness Scatter Plot

The `psa_scatter_plot()` function shows costs vs outcomes for all simulations.

#### Basic usage

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

#### Filtering to specific strategies

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  strategies = c("standard", "new_treatment")
)
```

#### Hiding mean points

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  show_means = FALSE
)
```

### Pairwise Incremental Scatter Plot

The `pairwise_psa_scatter_plot()` function shows incremental costs and outcomes relative to a comparator.

#### Basic incremental CE plane

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  comparator = "standard"
)
```

#### With WTP threshold line

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  comparator = "standard",
  wtp = 50000
)
```

## CEAC Plots

Cost-effectiveness acceptability curves (CEACs) show the probability of being cost-effective across willingness-to-pay (WTP) thresholds.

### Incremental CEAC

The `incremental_ceac_plot()` function shows the probability that each strategy has the highest net monetary benefit.

#### Basic usage

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

#### With cost-effectiveness frontier overlay

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  show_frontier = TRUE
)
```

#### Custom WTP range

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  wtp_range = c(0, 200000),
  wtp_step = 5000
)
```

### Pairwise CEAC

The `pairwise_ceac_plot()` function shows the probability that an intervention is cost-effective compared to a specific comparator.

#### Intervention perspective

```{r, fig.width=9, fig.height=6}
pairwise_ceac_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  intervention = "new_treatment"
)
```

#### Comparator perspective

```{r, fig.width=9, fig.height=6}
pairwise_ceac_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  comparator = "standard"
)
```

### CEAC Frontier Plot

The `incremental_ceac_frontier_plot()` function shows which strategy is optimal at each WTP threshold.

```{r, fig.width=9, fig.height=6}
incremental_ceac_frontier_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

## EVPI Plot

The expected value of perfect information (EVPI) represents the maximum amount a decision maker should pay to eliminate all parameter uncertainty.

### Basic EVPI plot

```{r, fig.width=9, fig.height=6}
evpi_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

### Custom WTP range

```{r, fig.width=9, fig.height=6}
evpi_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  wtp_range = c(0, 200000),
  wtp_step = 5000
)
```

## NMB Density Plot

The `nmb_density_plot()` function shows the distribution of net monetary benefit across simulations.

### Basic usage

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  wtp = 50000
)
```

### With mean lines

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  wtp = 50000,
  show_mean = TRUE
)
```

## PSA Tables

### PSA Summary Table

The `psa_summary_table()` function provides comprehensive summary statistics including means, standard deviations, and probability of cost-effectiveness at various WTP thresholds.

#### Basic usage

```{r}
psa_summary_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

#### With custom WTP thresholds for P(CE)

```{r}
psa_summary_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  pce_wtp = c(20000, 50000, 100000, 150000)
)
```

#### Filtering to specific strategies

```{r}
psa_summary_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  strategies = c("standard", "new_treatment")
)
```

### Incremental CEAC Table

The `incremental_ceac_table()` function presents CEAC probabilities at selected WTP thresholds.

#### Basic usage

```{r}
incremental_ceac_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

#### With optimal strategy highlighting

```{r}
incremental_ceac_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  show_optimal = TRUE
)
```

#### Custom WTP thresholds

```{r}
incremental_ceac_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  wtp_thresholds = c(20000, 50000, 100000, 150000, 200000)
)
```

### Pairwise CEAC Table

The `pairwise_ceac_table()` function shows pairwise cost-effectiveness probabilities.

#### Intervention perspective

```{r}
pairwise_ceac_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  intervention = "new_treatment"
)
```

#### Comparator perspective

```{r}
pairwise_ceac_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  comparator = "standard"
)
```

### EVPI Table

The `evpi_table()` function presents EVPI values at selected WTP thresholds.

```{r}
evpi_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs"
)
```

### CE Quadrant Table

The `ce_quadrant_table()` function shows the percentage of simulations falling into each quadrant of the cost-effectiveness plane relative to a comparator.

```{r}
ce_quadrant_table(
  psa_results,
  outcome_summary = "qalys",
  cost_summary = "costs",
  comparator = "Standard Care"
)
```

### PSA Parameters Table

The `psa_parameters_table()` function summarizes the distribution of sampled parameter values.

```{r}
psa_parameters_table(
  psa_results,
  variables = c("p_mild", "p_severe", "u_mild", "u_severe")
)
```

### Parameter Scatter Matrix

The `psa_parameter_scatter_matrix()` function creates a scatterplot matrix showing relationships between sampled parameters, including correlations, scatter plots, and density distributions.

```{r, fig.width=9, fig.height=9}
psa_parameter_scatter_matrix(
  psa_results,
  variables = c("p_mild", "p_severe", "u_mild", "u_severe"),
  strategies = "standard"
)
```

