---
title: "Probabilistic Sensitivity Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Probabilistic Sensitivity Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to configure, run, and visualize probabilistic sensitivity analysis (PSA) using openqaly. PSA simultaneously varies all uncertain parameters according to their probability distributions to characterize overall uncertainty in model results.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
# Skip on CI and CRAN (R CMD check sets _R_CHECK_PACKAGE_NAME_)
if (identical(Sys.getenv("CI"), "true") || nzchar(Sys.getenv("_R_CHECK_PACKAGE_NAME_"))) {
  message("Skipping vignette build on CI/CRAN")
  knitr::knit_exit()
}

library(openqaly)

# Build a simple Markov model with PSA sampling distributions
model <- define_model("markov") |>
  set_settings(
    n_cycles = 20,
    cycle_length = 1,
    cycle_length_unit = "years",
    discount_cost = 3.5,
    discount_outcomes = 3.5
  ) |>
  # States
  add_state("mild", initial_prob = 1) |>
  add_state("severe", initial_prob = 0) |>
  add_state("dead", initial_prob = 0) |>
  # Strategies
  add_strategy("seritinib", display_name = "Seritinib") |>
  add_strategy("volantor", display_name = "Volantor") |>
  add_strategy("cendralimab", display_name = "Cendralimab") |>
  # Groups
  add_group("low_risk", display_name = "Low Risk") |>
  # Base transition probabilities with univariate sampling
  add_variable("p_mild", 0.15, sampling = beta(bc, bc * 0.2)) |>
  add_variable("p_severe_base", 0.10, sampling = beta(bc, bc * 0.2)) |>
  add_variable("p_death_severe", 0.05, sampling = beta(bc, bc * 0.25)) |>
  add_variable("p_death_mild", 0.02) |>
  # Treatment effect on progression (strategy-specific)
  add_variable("treatment_effect", 1.0, strategy = "seritinib") |>
  add_variable("treatment_effect", 0.7, strategy = "volantor") |>
  add_variable("treatment_effect", 0.5, strategy = "cendralimab") |>
  # Effective progression probability
  add_variable("p_severe", p_severe_base * treatment_effect) |>
  # Costs with sampling
  add_variable("c_mild", 2500, sampling = gamma(bc, bc * 0.2)) |>
  add_variable("c_severe", 8000, sampling = gamma(bc, bc * 0.2)) |>
  add_variable("c_treatment", 3000, strategy = "seritinib",
               sampling = gamma(bc, bc * 0.15)) |>
  add_variable("c_treatment", 5000, strategy = "volantor",
               sampling = gamma(bc, bc * 0.15)) |>
  add_variable("c_treatment", 7000, strategy = "cendralimab",
               sampling = gamma(bc, bc * 0.15)) |>
  # Utilities (base values, sampled via multivariate)
  add_variable("u_mild", 0.85) |>
  add_variable("u_severe", 0.55) |>
  # Covariance table for correlated utilities
  # sd_mild=0.05, sd_severe=0.08, cor=0.6
  # cov(mild,mild) = 0.05^2 = 0.0025
  # cov(severe,severe) = 0.08^2 = 0.0064
  # cov(mild,severe) = 0.6 * 0.05 * 0.08 = 0.0024
  add_table("utility_cov", data.frame(
    u_mild = c(0.0025, 0.0024),
    u_severe = c(0.0024, 0.0064)
  )) |>
  # Multivariate sampling for correlated utilities
  add_multivariate_sampling(
    name = "utility_correlation",
    type = "mvnormal",
    variables = c("u_mild", "u_severe"),
    covariance = "utility_cov"
  ) |>
  # Transitions
  add_transition("mild", "severe", "p_severe") |>
  add_transition("mild", "dead", "p_death_mild") |>
  add_transition("mild", "mild", "1 - p_severe - p_death_mild") |>
  add_transition("severe", "dead", "p_death_severe") |>
  add_transition("severe", "severe", "1 - p_death_severe") |>
  add_transition("dead", "dead", "1") |>
  # Values
  add_value("cost", "c_mild + c_treatment", state = "mild", type = "cost") |>
  add_value("cost", "c_severe + c_treatment", state = "severe", type = "cost") |>
  add_value("cost", "0", state = "dead", type = "cost") |>
  add_value("qaly", "u_mild", state = "mild", type = "outcome") |>
  add_value("qaly", "u_severe", state = "severe", type = "outcome") |>
  add_value("qaly", "0", state = "dead", type = "outcome") |>
  # Summaries
  add_summary("qalys", "qaly", type = "outcome", display_name = "QALYs") |>
  add_summary("costs", "cost", type = "cost", display_name = "Costs")
```

## Configuring PSA Sampling Distributions

Before running PSA, you must specify probability distributions for uncertain parameters using the `sampling` argument in `add_variable()`.

### Basic usage with a normal distribution

```{r, eval = FALSE}
model <- model %>%
  add_variable("c_severe", 8000, sampling = normal(bc, 1000))
```

### Using the `bc` keyword for relative specifications

The `bc` keyword references the base case value, allowing relative distribution specifications.

```{r, eval = FALSE}
model <- model %>%
  add_variable("p_mild", 0.15, sampling = normal(bc, bc * 0.1))
```

### Different distribution types

Common distributions for health economic parameters:

```{r, eval = FALSE}
# Beta distribution for probabilities (constrained to 0-1)
model <- model %>%
  add_variable("p_death", 0.02, sampling = beta(bc, bc * 0.2))

# Lognormal distribution for costs (positive values only)
model <- model %>%
  add_variable("c_treatment", 5000, sampling = lognormal(bc, bc * 0.3))

# Gamma distribution for resource use
model <- model %>%
  add_variable("n_visits", 12, sampling = gamma(bc, bc * 0.25))
```

### Multivariate sampling for correlated parameters

Use `add_multivariate_sampling()` when parameters should be sampled together. Three distribution types are supported: multivariate normal, Dirichlet, and multinomial.

#### Multivariate normal with covariance from a model table

Store the covariance matrix as a model table (e.g., from `vcov()` of a fitted model) and reference it by name. Means are taken from the variable base case values automatically.

```{r, eval = FALSE}
# Example: survival model parameters from flexsurv
# fit <- flexsurvreg(Surv(time, status) ~ 1, data = dat, dist = "gengamma")
# vcov(fit) produces a 3x3 covariance matrix for mu, sigma, Q
model <- model |>
  add_variable("mu", 2.1) |>
  add_variable("sigma", 0.8) |>
  add_variable("Q", 0.5) |>
  add_table("surv_vcov", as.data.frame(vcov(fit))) |>
  add_multivariate_sampling(
    name = "survival_params",
    type = "mvnormal",
    variables = c("mu", "sigma", "Q"),
    covariance = "surv_vcov"
  )
```

#### Dirichlet for transition probabilities

Dirichlet sampling ensures probabilities sum to 1. Specify `n` (effective sample size) and alpha values are derived as `n * base_case_value`.

```{r, eval = FALSE}
model <- model |>
  add_variable("p_stable", 0.7) |>
  add_variable("p_prog", 0.2) |>
  add_variable("p_death", 0.1) |>
  add_multivariate_sampling(
    name = "sick_transitions",
    type = "dirichlet",
    variables = c("p_stable", "p_prog", "p_death"),
    n = 100  # alpha = c(70, 20, 10)
  )
```

You can also pass an explicit `alpha` vector instead of `n`:

```{r, eval = FALSE}
model <- model |>
  add_multivariate_sampling(
    name = "sick_transitions",
    type = "dirichlet",
    variables = c("p_stable", "p_prog", "p_death"),
    alpha = c(70, 20, 10)
  )
```

#### Multinomial for categorical outcomes

Multinomial sampling produces one-hot encoded outcomes. Probabilities are taken from variable base case values.

```{r, eval = FALSE}
model <- model |>
  add_variable("init_s1", 0.5) |>
  add_variable("init_s2", 0.3) |>
  add_variable("init_s3", 0.2) |>
  add_multivariate_sampling(
    name = "initial_states",
    type = "multinomial",
    variables = c("init_s1", "init_s2", "init_s3"),
    size = 1
  )
```

#### Strategy-specific multivariate sampling

Use the `strategy` parameter to restrict sampling to a specific strategy:

```{r, eval = FALSE}
model <- model |>
  add_multivariate_sampling(
    name = "intervention_costs",
    type = "mvnormal",
    variables = c("drug_cost", "admin_cost"),
    strategy = "intervention",
    covariance = "cost_cov_table"
  )
```

## Running PSA

```{r, include = FALSE}
# Run PSA for the examples
psa_results <- run_psa(model, n_sim = 1000, seed = 123)
```

The `run_psa()` function executes the model for each Monte Carlo simulation, sampling parameters from their specified distributions.

```{r, eval = FALSE}
psa_results <- run_psa(model, n_sim = 1000, seed = 123)
```

## Outcome Results

### Outcome Density Plot

The `outcome_density_plot()` function creates density plots of outcome values from PSA simulations. It can display either absolute outcome distributions per strategy or distributions of outcome differences between intervention/comparator pairs.

#### Basic usage (absolute mode)

Display outcome distributions for all strategies.

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab")
)
```

#### Filtering to specific strategies

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor")
)
```

#### Incremental mode (comparator perspective)

Show outcome differences relative to a comparator.

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  comparators = "seritinib"
)
```

#### Incremental mode (intervention perspective)

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  interventions = "volantor"
)
```

#### Multiple comparisons

Compare multiple interventions against a common comparator.

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib"
)
```

#### Hiding mean lines

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab"),
  show_mean = FALSE
)
```

#### Undiscounted outcomes

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab"),
  discounted = FALSE
)
```

#### Results by population subgroup

Use `groups = "all_groups"` to display results for each population subgroup separately.

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab"),
  groups = "all_groups"
)
```

Use `groups = "all"` to include both the overall population and all subgroups.

```{r, fig.width=9, fig.height=6}
outcome_density_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab"),
  groups = "all"
)
```

### Outcome Ridgeline Plot

The `outcome_ridgeline_plot()` function creates ridgeline plots as an alternative to density plots. All strategies are displayed on the y-axis with distributions along the x-axis, making it easy to compare across strategies in a single panel.

#### Basic usage (absolute mode)

```{r, fig.width=9, fig.height=5}
outcome_ridgeline_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab")
)
```

#### Incremental mode

```{r, fig.width=9, fig.height=5}
outcome_ridgeline_plot(
  psa_results,
  outcome = "qalys",
  comparators = "seritinib"
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=5}
outcome_ridgeline_plot(
  psa_results,
  outcome = "qalys",
  strategies = c("seritinib", "volantor", "cendralimab"),
  groups = "all_groups"
)
```

### Cost Density Plot

The `cost_density_plot()` function creates density plots of cost values from PSA simulations with currency formatting.

#### Basic usage (absolute mode)

```{r, fig.width=9, fig.height=6}
cost_density_plot(
  psa_results,
  outcome = "costs",
  strategies = c("seritinib", "volantor", "cendralimab")
)
```

#### Incremental mode

```{r, fig.width=9, fig.height=6}
cost_density_plot(
  psa_results,
  outcome = "costs",
  comparators = "seritinib"
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=6}
cost_density_plot(
  psa_results,
  outcome = "costs",
  strategies = c("seritinib", "volantor", "cendralimab"),
  groups = "all_groups"
)
```

### Cost Ridgeline Plot

The `cost_ridgeline_plot()` function creates ridgeline plots of cost distributions with currency formatting.

#### Basic usage (absolute mode)

```{r, fig.width=9, fig.height=5}
cost_ridgeline_plot(
  psa_results,
  outcome = "costs",
  strategies = c("seritinib", "volantor", "cendralimab")
)
```

#### Incremental mode

```{r, fig.width=9, fig.height=5}
cost_ridgeline_plot(
  psa_results,
  outcome = "costs",
  comparators = "seritinib"
)
```

### PSA Outcomes Table

The `psa_outcomes_table()` function displays summary statistics for outcome values from PSA simulations, including mean, SD, median (IQR), and range.

#### Basic usage

```{r}
psa_outcomes_table(
  psa_results,
  outcome = "qalys"
)
```

#### Outcome differences relative to comparator

```{r}
psa_outcomes_table(
  psa_results,
  outcome = "qalys",
  comparators = "seritinib"
)
```

#### Undiscounted outcomes

```{r}
psa_outcomes_table(
  psa_results,
  outcome = "qalys",
  discounted = FALSE
)
```

#### Results by population subgroup

```{r}
psa_outcomes_table(
  psa_results,
  outcome = "qalys",
  groups = "all_groups"
)
```

### PSA Costs Table

The `psa_costs_table()` function displays summary statistics for cost values from PSA simulations with currency formatting.

#### Basic usage

```{r}
psa_costs_table(
  psa_results,
  outcome = "costs"
)
```

#### Differences relative to comparator

```{r}
psa_costs_table(
  psa_results,
  outcome = "costs",
  comparators = "seritinib"
)
```

#### Results by population subgroup

```{r}
psa_costs_table(
  psa_results,
  outcome = "costs",
  groups = "all_groups"
)
```

## Cost-Effectiveness Results

### PSA Summary Table

The `psa_summary_table()` function provides comprehensive summary statistics including means, standard deviations, and probability of cost-effectiveness at various WTP thresholds.

#### Basic usage

```{r}
psa_summary_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### With custom WTP thresholds for P(CE)

```{r}
psa_summary_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  pce_wtp = c(20000, 50000, 100000, 150000)
)
```

#### Filtering to specific strategies

```{r}
psa_summary_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  strategies = c("seritinib", "volantor")
)
```

#### Results by population subgroup

```{r}
psa_summary_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

### CE Scatter Plot

Scatter plots visualize simulation results on the cost-effectiveness plane.

The `psa_scatter_plot()` function shows costs vs outcomes for all simulations.

#### Basic usage

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### Filtering to specific strategies

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  strategies = c("seritinib", "volantor")
)
```

#### Hiding mean points

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  show_means = FALSE
)
```

#### Results by population subgroup

Use `groups = "all_groups"` to display results for each population subgroup separately.

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

Use `groups = "all"` to include both the overall population and all subgroups.

```{r, fig.width=9, fig.height=6}
psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all"
)
```

### Pairwise Incremental Scatter Plot

The `pairwise_psa_scatter_plot()` function shows incremental costs and outcomes relative to a comparator.

#### Basic incremental CE plane

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparators = "seritinib"
)
```

#### With WTP threshold line

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparators = "seritinib",
  wtp = 50000
)
```

#### Multiple interventions

Compare multiple interventions against a common comparator.

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib",
  wtp = 50000
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=6}
pairwise_psa_scatter_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparators = "seritinib",
  groups = "all_groups"
)
```

### CE Quadrant Table

The `ce_quadrant_table()` function shows the percentage of simulations falling into each quadrant of the cost-effectiveness plane relative to a comparator.

#### Basic usage

```{r}
ce_quadrant_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparator = "Seritinib"
)
```

#### Results by population subgroup

```{r}
ce_quadrant_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparator = "Seritinib",
  groups = "all_groups"
)
```

## Cost-Effectiveness Acceptability Curves (CEAC)

Cost-effectiveness acceptability curves (CEACs) show the probability of being cost-effective across willingness-to-pay (WTP) thresholds.

### Incremental CEAC Plot

The `incremental_ceac_plot()` function shows the probability that each strategy has the highest net monetary benefit.

#### Basic usage

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### Custom WTP range

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp_range = c(0, 200000),
  wtp_step = 5000
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=6}
incremental_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

### Incremental CEAC Table

The `incremental_ceac_table()` function presents CEAC probabilities at selected WTP thresholds.

#### Basic usage

```{r}
incremental_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### Custom WTP thresholds

```{r}
incremental_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp_thresholds = c(20000, 50000, 100000, 150000, 200000)
)
```

#### Results by population subgroup

```{r}
incremental_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

### Pairwise CEAC Plot

The `pairwise_ceac_plot()` function shows the probability that an intervention is cost-effective compared to a specific comparator.

#### Intervention perspective

```{r, fig.width=9, fig.height=6}
pairwise_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  interventions = "volantor"
)
```

#### Comparator perspective

```{r, fig.width=9, fig.height=6}
pairwise_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparators = "seritinib"
)
```

#### Multiple comparisons

Compare multiple interventions against a common comparator.

```{r, fig.width=9, fig.height=6}
pairwise_ceac_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib"
)
```

### Pairwise CEAC Table

The `pairwise_ceac_table()` function shows pairwise cost-effectiveness probabilities.

#### Intervention perspective

```{r}
pairwise_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  interventions = "volantor"
)
```

#### Comparator perspective

```{r}
pairwise_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  comparators = "seritinib"
)
```

#### Multiple comparisons

```{r}
pairwise_ceac_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib"
)
```

## Net Monetary Benefit (NMB)

### Incremental NMB Density Plot

The `nmb_density_plot()` function shows the distribution of incremental net monetary benefit across simulations relative to a comparator strategy.

#### Basic usage (comparator mode)

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib"
)
```

#### Hiding mean lines

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib",
  show_mean = FALSE
)
```

#### Intervention mode

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 100000,
  interventions = "volantor"
)
```

#### Multiple comparisons

Compare multiple interventions against a common comparator.

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib"
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=6}
nmb_density_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib",
  groups = "all_groups"
)
```

### Incremental NMB Ridgeline Plot

The `nmb_ridgeline_plot()` function creates ridgeline plots of incremental NMB distributions. Comparisons are displayed on the y-axis with a vertical dashed line at zero indicating the cost-effectiveness threshold.

#### Basic usage (comparator mode)

```{r, fig.width=9, fig.height=5}
nmb_ridgeline_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib"
)
```

#### Multiple comparisons

```{r, fig.width=9, fig.height=5}
nmb_ridgeline_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  interventions = c("volantor", "cendralimab"),
  comparators = "seritinib"
)
```

### PSA Incremental NMB Table

The `psa_nmb_table()` function displays summary statistics for incremental net monetary benefit from PSA simulations, including mean, SD, median (IQR), range, and probability of positive NMB.

#### Comparator perspective

```{r}
psa_nmb_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib"
)
```

#### Intervention perspective

```{r}
psa_nmb_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 100000,
  interventions = "volantor"
)
```

#### Results by population subgroup

```{r}
psa_nmb_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp = 50000,
  comparators = "seritinib",
  groups = "all_groups"
)
```

## Expected Value of Perfect Information (EVPI)

The expected value of perfect information (EVPI) represents the maximum amount a decision maker should pay to eliminate all parameter uncertainty.

### EVPI Plot

#### Basic usage

```{r, fig.width=9, fig.height=6}
evpi_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### Custom WTP range

```{r, fig.width=9, fig.height=6}
evpi_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  wtp_range = c(0, 200000),
  wtp_step = 5000
)
```

#### Results by population subgroup

```{r, fig.width=9, fig.height=6}
evpi_plot(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

### EVPI Table

The `evpi_table()` function presents EVPI values at selected WTP thresholds.

#### Basic usage

```{r}
evpi_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs"
)
```

#### Results by population subgroup

```{r}
evpi_table(
  psa_results,
  health_outcome = "qalys",
  cost_outcome = "costs",
  groups = "all_groups"
)
```

## Parameter Diagnostics

### PSA Parameters Table

The `psa_parameters_table()` function summarizes the distribution of sampled parameter values.

```{r}
psa_parameters_table(
  psa_results,
  variables = c("p_mild", "p_severe", "u_mild", "u_severe")
)
```

### Parameter Scatter Matrix

The `psa_parameter_scatter_matrix()` function creates a scatterplot matrix showing relationships between sampled parameters, including correlations, scatter plots, and density distributions.

```{r, fig.width=9, fig.height=9}
psa_parameter_scatter_matrix(
  psa_results,
  variables = c("p_mild", "p_severe", "u_mild", "u_severe"),
  strategies = "seritinib",
  group = "low_risk"
)
```
